\section{Introduction}
\label{sec:einleitung}

Large Language Models (LLMs) have become a component of software that reads
emails, summarises documents, retrieves web content, and executes actions on
behalf of users. This integration makes them useful, but also exposes them to
inputs controlled by parties other than the application operator.

The threat considered in this paper is \emph{prompt injection}. An attacker
places a malicious instruction in content processed by the model, and the model
may follow it instead of the trusted task
\parencites{perez2022ignore}{greshake2023indirect}. The resulting behaviour can
redirect an answer, disclose protected information, or influence a connected
tool. The 2025 Open Worldwide Application Security Project (OWASP) Top 10 for
LLM Applications lists prompt injection as LLM01, reflecting its importance as
a system-security risk~\parencite{owasp2025llm}.

Applications can encode developer, user, and tool messages with distinct roles
or delimiters. These distinctions affect model behaviour, but they are still
interpreted by the model and should not be treated as access control. Unlike a
prepared Structured Query Language (SQL) statement, a prompt format does not by
itself give natural-language text a type that prevents it from influencing
generation. The relevant weakness
is therefore not that all input is literally identical, but that no reliably
enforced provenance or authority boundary can be assumed from model behaviour
alone
\parencites{liu2023promptinjection}{wallace2024instruction}.

Research addresses this problem at several layers. Model-level approaches train
the LLM to prioritise trusted instructions; application-level approaches detect,
label, or validate data and outputs; system-level controls restrict tools,
credentials, and information flows. This paper retains that broader
system-security perspective, but its detailed empirical analysis is deliberately
narrower. StruQ and SecAlign are used as case studies of learned
instruction--data separation, while other defence classes are organised and
discussed without claiming equally detailed experimental coverage.

\medskip
\noindent\textbf{Research Question} \quad
\textit{How effectively do learned instruction--data separation approaches
mitigate prompt injection under the evaluated attacker models, and what
system-level controls are required when they do not provide an independently
enforced security boundary?}

\medskip
This work is a scoped narrative review rather than a systematic review. It
considers the sources in the project bibliography available up to 19 July 2026.
StruQ and SecAlign were selected because they implement closely related learned
separation strategies, publish security and utility results, and are examined by
subsequent broader and defence-adaptive evaluations. Those later studies were
selected where they test the case-study defences under changed prompts,
attacker knowledge, optimisation methods, or transfer settings. Primary sources
are preferred; recent preprints are included because the technical evidence is
developing rapidly. No original experiments or independent replications were
conducted for this paper.

Defence effectiveness is assessed across attack success rate, attacker
adaptivity, attacker access and knowledge, model and task coverage, utility
impact, transferability, deployment assumptions, and the location and nature of
enforcement. Percentages from different papers are not treated as directly
comparable when their models, attacks, optimisation budgets, success rules, or
datasets differ. In particular, behavioural robustness is probabilistic, whereas
an external policy mechanism may be deterministic only under its stated system
assumptions.

The thesis of this paper is that learned separation of trusted instructions from
untrusted data can substantially reduce prompt-injection success under evaluated
conditions, but it is a mitigation rather than independently enforced
prevention. The absence of a demonstrated deterministic guarantee is not an
impossibility proof for all model-level defences. Secure deployment must
therefore combine model-level protection with system-level containment and
defence in depth.

The remainder of this paper is organised as follows.
Section~\ref{sec:llms} introduces LLMs and prompting, followed by direct and
indirect prompt injection in Section~\ref{sec:attacks}.
Section~\ref{sec:threat} defines the system, attacker, and security assumptions.
Section~\ref{sec:struq} positions the defence landscape and examines StruQ and
SecAlign. Section~\ref{sec:limits} compares their evidence, adaptive attacks,
and security--utility trade-offs. Section~\ref{sec:praxis} derives practical
deployment implications before Section~\ref{sec:fazit} concludes.
