\section{LLMs und Prompting}
\label{sec:llms}

Before the security implications of prompt injection attacks can be discussed and how they can be averted, it is important to understand the technology they use, namely LLMs and prompting.\\
LLMs are a type of artificial intelligence model (AI) that is trained on large amounts of text data to understand and generate human-like language. 
They are designed to predict the next pieces of a word, so called tokens, in a sequence.
This allows them to generate coherent and contextually relevant text based on the input they receive.\\
This input is called a prompt. It is split up into two parts: the user prompt and the system prompt.
The user prompt is a set of instructions given to the AI model by the user, telling the LLM what it should do.
The user prompt is then sent to the server which the AI model runs on and embedded into the system prompt.
The system prompt is a set of instructions set by the developers of the LLM, which is used to guide the behavior of the model and ensure that it generates appropriate responses.
It can, among other things, tell the model how to respond to different kinds of user prompts, whilst also ensuring, that the model does not generate harmful or inappropriate content.\\
From this architecture, it becomes clear that the user prompt is a potential attack vector, as user prompts are embedded into the system prompt, combining them into a single prompt that is then processed by the LLM.
If an attacker is able to manipulate their prompt in such a way that it overrides the instructions from the system prompt, they can potentially cause the model to generate inappropriate or confidential information\parencite{liu2023promptinjection}.\\
This is where prompt injection attacks come into play, which we will discuss in more detail in the next section.\\