\section{Threat Model}
\label{sec:threat}

A threat model states which components and inputs are trusted, which access an
attacker may obtain, and which security properties a defence should preserve.
This is necessary because prompt-injection evaluations make substantially
different assumptions. Resistance to malicious text in a document does not
imply resistance to an attacker with model gradients, defence details, or
fine-tuning checkpoints.

This paper considers text-based prompt injection against LLM-integrated
applications. It distinguishes the target system, the access and knowledge used
during payload construction, the channel used for payload delivery, and the
security consequences of a successful attack.

\subsection{System Model}
\label{subsec:system-model}

The considered system contains five principal components:

\begin{enumerate}
    \item a trusted instruction and policy source controlled by the application
          operator;
    \item untrusted input channels, including direct user input and external
          data sources;
    \item an application layer that assembles context, invokes the model, and
          applies release or action policies outside it;
    \item an LLM that generates an output or proposes a tool call; and
    \item connected tools or services that may execute a proposed action,
          subject to application-level authorisation checks.
\end{enumerate}

The system also involves distinct human or organisational principals: the
attacker, the legitimate user, the application operator, the owner of processed
data, and the provider of a connected tool or service. One party may occupy
several roles, but their authority is not interchangeable. In particular,
attacker-controlled content does not acquire the user's or operator's authority
merely because the model processes it in a session that can access their data or
credentials
\parencites{greshake2023indirect}{debenedetti2024agentdojo}.

The trusted instruction defines the intended task, such as summarising a
document or extracting information from an email. Untrusted data may originate
from a user, website, retrieved document, email, uploaded file, API response, or
tool output. The application may also contain RAG pipelines, prompt templates,
memory, and interfaces to tools. These components determine which information
reaches the model and which consequences its output may have.

Figure~\ref{fig:threat-model} shows the principal trust transitions. Trusted and
untrusted inputs both reach the application context. Generated output should be
released only to an authorised recipient and for a permitted purpose; egress
checks may additionally constrain protected data, destinations, or active
content. Proposed external actions pass through a separate policy and
authorisation component. These controls represent the system-level protection
recommended later; they are not assumed to make the model itself resistant to
injection.

\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}[
        box/.style={draw, rounded corners, align=center, text width=2.25cm,
                    minimum height=0.9cm},
        trusted/.style={box, thick},
        untrusted/.style={box, dashed},
        flow/.style={-{Latex[length=2.2mm]}, thick}
    ]
        \node[trusted] (instruction) at (1.3,1.5)
            {Trusted instruction\\and policy};
        \node[untrusted] (external) at (1.3,-1.5)
            {Untrusted direct input\\or external content};
        \node[box] (application) at (4.6,0)
            {Application\\context assembly};
        \node[box] (llm) at (7.8,0)
            {Large Language\\Model};
        \node[box] (output) at (10.9,1.5)
            {Generated response\\to user};
        \node[trusted, text width=2.5cm] (policy) at (10.9,-1.5)
            {Policy and\\authorisation gate};
        \node[box, text width=1.8cm] (tool) at (13.8,-1.5)
            {Tool or\\service};

        \draw[flow] (instruction) -- (application);
        \draw[flow] (external) -- (application);
        \draw[flow] (application) -- (llm);
        \draw[flow] (llm) -- (output);
        \draw[flow] (llm) -- (policy);
        \draw[flow] (policy) -- (tool);
    \end{tikzpicture}
    \caption[Threat model for an LLM-integrated application]{Simplified threat
    model for an LLM-integrated application. Trusted instructions and
    untrusted content are assembled into the model context. Generated responses
    may be returned to the user, while proposed tool actions pass through a
    policy and authorisation gate. Original schematic based on
    \textcite{greshake2023indirect} and
    \textcite{debenedetti2024agentdojo}.}
    \label{fig:threat-model}
\end{figure}

A read-only application may be limited to a manipulated response or disclosure
of information. An agentic application may allow the same model-level failure
to influence messages, file operations, or remote services. The security impact
therefore depends on the authority and protected data exposed by the system, not
only on whether the model followed an injected string
\parencite{debenedetti2024agentdojo}.

\subsection{Attacker Model}
\label{subsec:attacker-model}

The attacker controls at least one untrusted input channel. In a direct attack,
the attacker submits malicious input through a user-facing interface. In an
indirect attack, the attacker controls content later processed by the
application, for example by publishing a webpage, sending an email, supplying a
document, or influencing an API or tool response
\parencite{greshake2023indirect}.

Indirect delivery does not require the attacker to query the target model. It is
sufficient that the application later includes the malicious content in the
model context. This delivery condition remains separate from any model access
used to construct the payload.

Attacker capability is recorded along five dimensions:

\begin{enumerate}
    \item query access to the target or a surrogate model;
    \item access to target weights, gradients, activations, or checkpoints;
    \item knowledge of the trusted instruction and surrounding context;
    \item knowledge of the deployed defence and application format; and
    \item the direct or indirect channel available for delivery.
\end{enumerate}

These dimensions prevent black-, grey-, and white-box terminology from being
used as a shorthand for all attacker knowledge.

\subsubsection{Black-Box Model Access}

A black-box attacker has no access to target-model weights, gradients, internal
activations, or checkpoints. The attacker may still query the model or observe
externally visible behaviour. Black-box model access does not imply weak system
knowledge: an attacker may know the prompt template or deployed defence without
seeing model internals.

Handcrafted direct and indirect injections commonly fit this access model. A
black-box attacker may nevertheless adapt a payload through repeated queries if
the target exposes sufficient feedback.

\subsubsection{Grey-Box Model Access}

A grey-box attacker has partial internal information or access to a related
surrogate, but not the complete target internals required to compute exact
gradients for the deployed model. Knowledge of the model family, delimiter
format, or defence mechanism should be reported separately; none is sufficient
on its own to define grey-box access.

This distinction is relevant to transfer attacks. A suffix may be constructed
on an accessible surrogate and then tested against a related target. Transfer
must be demonstrated empirically rather than assumed.

\subsubsection{White-Box Model Access}

A white-box attacker has the internal information required by the chosen
optimisation method. GCG uses model gradients, Checkpoint-GCG additionally uses
intermediate fine-tuning checkpoints, and ASTRA uses weights, gradients, and
internal attention information
\parencites{zou2023universal}{yang2025checkpointgcg}{pandya2025astra}.

This access is less representative of a direct attack against a closed
commercial model. It remains relevant for open-weight deployments and for
auditing whether robustness persists when an attacker adapts to the defence.
The realism of payload delivery and the strength of payload construction should
therefore be assessed separately.

\subsubsection{Context Knowledge, Defence Knowledge, and Delivery}

Context knowledge records whether the attacker knows the trusted instruction,
prompt template, private conversation history, and other text in the target
context. Defence knowledge records whether the attacker knows filters,
delimiters, defensive training, or policy checks.

ASTRA illustrates why these dimensions cannot be collapsed. Its
\emph{strong-knowledge} and \emph{weak-knowledge} settings both assume an
adaptive white-box attacker with model and defence access. Strong knowledge
additionally reveals the preceding context; weak knowledge withholds that
context. Neither term is synonymous with black- or grey-box model access
\parencite{pandya2025astra}.

Delivery is independent again. A payload constructed with white-box access may
be placed in a document for indirect delivery. Conversely, a direct attacker
may have black-box model access but detailed knowledge of the deployed defence.

\subsection{Security Goals}
\label{subsec:security-goals}

A prompt-injection defence should preserve three principal properties.

\paragraph{Instruction integrity.}

Untrusted content may and often must influence the task-relevant information
processed by the model. It must not thereby acquire authority to alter security
policy, authorise actions, redirect protected information, or override trusted
control instructions. Instructions quoted inside a document, email, website, or
tool output may be analysed as data without gaining the authority of the trusted
instruction channel
\parencites{chen2025struq}{chen2025secalign}.

\paragraph{Confidentiality of protected context.}

Attacker-controlled content should not cause the application to disclose system
instructions, private user data, retrieved secrets, or other information to a
principal that is not authorised to receive it.

\paragraph{Action integrity.}

In systems with tool access, attacker-controlled content should not cause an
unauthorised operation or manipulate security-relevant parameters. Examples
include sending a message, modifying a file, selecting an attacker-controlled
recipient, or transmitting protected data
\parencites{fspib2025}{debenedetti2024agentdojo}.

These goals distinguish a security violation from harmless exposure to
malicious text. An attack succeeds in the security sense when it violates
instruction integrity, confidentiality, or action integrity.

The target-output Attack Success Rate (ASR) used in many model-level
evaluations is mainly a proxy for instruction-following compromise. It does not
by itself show that protected information was disclosed or that an unauthorised
tool action was executed. System-level evaluations must therefore measure
goal-specific outcomes in addition to target-string compliance
\parencites{jia2025critical}{debenedetti2024agentdojo}.

\subsection{Assumptions and Scope}
\label{subsec:threat-scope}

The trusted instruction source, application policy, and defence implementation
are assumed not to have been directly compromised. If an attacker can modify the
trusted prompt, replace the model, alter application code, or change the
authorisation policy, the problem is no longer limited to prompt injection.

\emph{In scope} are runtime, text-based prompt injections delivered through a
direct malicious input or through content processed by an otherwise
uncompromised application. This includes malicious instructions in retrieved
documents, emails, webpages, API responses, and outputs returned by connected
tools. Untrusted tool output is in scope even when the tool implementation is
not compromised.

\emph{Partially in scope} are social engineering and availability consequences.
Social engineering is considered only as a residual risk of human approval, not
as an empirically evaluated attack class. FSPIB treats response refusal as an
information-level threat, whereas resource exhaustion is an availability
impact. Neither denial-of-service resilience nor availability is evaluated as a
security goal here
\parencite{fspib2025}.

\emph{Out of scope} are multimodal injection, training-data poisoning, direct
compromise of a retriever or index, application or infrastructure compromise,
malicious tool implementations, insider threats, model-provider compromise, and
cross-session or memory-based persistence. Stored untrusted text can still
deliver an ordinary indirect injection when reloaded, but attacks that corrupt
the memory mechanism itself are not analysed.
