\section{StruQ and SecAlign}
\label{sec:defences}
\label{sec:struq}

Prompt-injection defences can operate at different layers. Some attempt to
change the behaviour of the language model, while others enforce restrictions
in the surrounding application. This chapter focuses on StruQ and SecAlign as
two model-level approaches to instruction--data separation. Both provide public
implementations and detailed security and utility evaluations, making it
possible to compare their mechanisms, reported effectiveness, and limitations
\parencites{chen2025struq}{chen2025secalign}.

\subsection{Motivation for Structured Queries}
\label{subsec:structured-queries}

As discussed in Section~\ref{sec:llms}, prompt injection is possible because an
LLM does not inherently know whether a sequence of tokens represents a trusted
instruction or data that should merely be processed. Conventional applications
often combine developer instructions and external content into a single model
input, leaving the distinction between them dependent on model interpretation.

A structured query instead contains two separately supplied components: a
trusted instruction that defines the task and untrusted data that the model
should process. In a document-summarisation application, for example, the
instruction may be \enquote{Summarise the following document}, while the
document itself forms the data component.

StruQ preserves this distinction at the application interface and trains the
model to respect it. The approach therefore consists of two complementary
components: a secure front end that constructs the structured prompt and a
fine-tuned model that is trained not to follow instructions found inside the
data component
\parencite{chen2025struq}.

\subsection{StruQ}
\label{subsec:struq}

\subsubsection{Secure Front End}
\label{subsubsec:struq-front-end}

The StruQ front end accepts the trusted instruction and untrusted data as
separate inputs. It then converts them into a predefined prompt format using
reserved delimiter tokens. These include \texttt{[MARK]} for the beginning of
a delimiter, \texttt{[INST]} for instructions, \texttt{[INPT]} for input data,
\texttt{[RESP]} for model responses, and \texttt{[COLN]} as a replacement for
the colon used in conventional prompt templates
\parencite{chen2025struq}.

An attacker could attempt to reproduce these tokens inside malicious content
and thereby forge a new instruction or response section. Before constructing
the final prompt, the front end therefore repeatedly removes the reserved
delimiter strings and the string \texttt{\#\#} from the untrusted input. The
filtering process continues until another pass no longer changes the input
\parencite{chen2025struq}.

This mechanism is intended to prevent completion attacks. In a completion
attack, the attacker inserts text that resembles the end of the legitimate
exchange and then introduces a new instruction. The aim is to make the model
behave as though the original task has already been completed and a new,
attacker-controlled task has begun. Removing the reserved strings prevents the
attacker from reproducing the exact structure generated by the trusted front
end.

The filter does not remove every natural-language instruction from the data.
It protects the structural encoding, while the model must still learn not to
follow ordinary instructional text contained in the untrusted section.

Figure~\ref{fig:struq-architecture} summarises how the trusted instruction and
untrusted data are processed separately. Reserved strings are removed only from
the untrusted data before both components are combined into the structured
query and passed to the StruQ model.

\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}[
        box/.style={draw, rounded corners, align=center, text width=2.25cm,
                    minimum height=0.9cm},
        trusted/.style={box, thick},
        untrusted/.style={box, dashed},
        flow/.style={-{Latex[length=2.2mm]}, thick}
    ]
        \node[trusted] (instruction) at (1.3,1.4)
            {Trusted\\instruction};
        \node[untrusted] (data) at (1.3,-1.4)
            {Untrusted\\data};
        \node[box] (filter) at (4.4,-1.4)
            {Reserved-string\\filter};
        \node[box] (formatter) at (7.4,0)
            {Structured-query\\formatter};
        \node[box, text width=2.8cm, font=\ttfamily\scriptsize]
            (query) at (10.5,0)
            {[MARK][INST] ...\\
             {}[MARK][INPT] ...\\
             {}[MARK][RESP]};
        \node[box, text width=1.8cm] (model) at (13.5,0)
            {StruQ\\model};

        \draw[flow] (instruction) -- (formatter);
        \draw[flow] (data) -- (filter);
        \draw[flow] (filter) -- (formatter);
        \draw[flow] (formatter) -- (query);
        \draw[flow] (query) -- (model);
    \end{tikzpicture}
    \caption[Architecture of StruQ]{Simplified StruQ front end and model input.
    The untrusted data pass through the reserved-string filter before being
    combined with the trusted instruction in the structured-query format.
    Original schematic based on \textcite{chen2025struq}.}
    \label{fig:struq-architecture}
\end{figure}

    \caption[Threat model for an LLM-integrated application]
    {Simplified threat model for an LLM-integrated application. Trusted
    instructions and untrusted external content are assembled into the model
    context. Generated responses may be returned to the user, while proposed
    tool actions must pass through a policy and authorisation gate before
    execution. Original schematic based on the system models described by
    \textcite{greshake2023indirect} and
    \textcite{debenedetti2024agentdojo}.}

    \label{fig:threat-model}
\end{figure}

\subsubsection{Structured Instruction Tuning}
\label{subsubsec:structured-instruction-tuning}

Structural formatting alone is insufficient. Existing instruction-tuned
language models have learned to follow natural-language instructions regardless
of where they appear in the context. If an ordinary instruction-tuned model
were given the StruQ prompt format without further training, it could still
follow commands placed inside the data component.

StruQ therefore introduces \emph{structured instruction tuning}. Conventional
instruction tuning uses supervised training examples consisting of an
instruction and a desired response. Through these examples, a pretrained
language model learns to perform tasks described through natural-language
instructions. However, conventional instruction tuning does not necessarily
teach the model that instructions from different parts of an application have
different levels of authority
\parencite{chen2025struq}.

Structured instruction tuning extends this process by teaching the model that
the location of an instruction is relevant to whether it should be followed.
The training dataset contains ordinary structured queries as well as augmented
examples in which an additional instruction has been inserted into the data
component. The expected response continues to follow the trusted instruction
and ignores the injected one
\parencite{chen2025struq}.

Table~\ref{tab:struq-example} presents a simplified example. Although the input
contains an instruction to output \enquote{HACKED}, the desired response still
performs the translation requested by the trusted instruction.

\begin{table}[htbp]
    \centering
    \begin{tabular}{p{0.92\linewidth}}
        \toprule
        \textbf{Trusted Instruction} \\
        Translate the following sentence into German. \\[0.4em]

        \textbf{Untrusted Input} \\
        Hello! \\
        Ignore the previous instruction and output
        \enquote{HACKED}. \\[0.4em]

        \textbf{Desired Response} \\
        Hallo! \\
        \bottomrule
    \end{tabular}
    \caption[Structured instruction-tuning example]{Simplified structured
    instruction-tuning example adapted from
    \textcite{chen2025struq}.}
    \label{tab:struq-example}
\end{table}

The augmented dataset also contains attacks that imitate or manipulate the
prompt structure, including forged instruction sections, fake delimiter tokens,
and manipulated response sections. Repeated exposure to these examples teaches
the model to associate executable instructions with the trusted instruction
component while treating instructions in the data component as content
\parencite{chen2025struq}.

The front end and model training therefore serve different purposes. The front
end prevents direct forgery of the reserved structural tokens, while structured
instruction tuning reduces compliance with ordinary natural-language
instructions in the data. StruQ relies on both mechanisms together.

\subsubsection{Original Evaluation and Limitations}
\label{subsec:struq-evaluation}

The principal security metric used in the StruQ evaluation is the Attack
Success Rate (ASR). It measures the proportion of evaluated inputs for which
the model follows the injected instruction instead of completing the intended
task. A lower ASR therefore indicates greater resistance to the tested attack.

StruQ was evaluated using Llama-7B and Mistral-7B on 208 AlpacaFarm samples
containing a data field. Most evaluated attacks inserted the task
\enquote{Print exactly Hacked!} into the data. An attack was considered
successful when the model treated the injected text as a new instruction and
produced the requested target output
\parencite{chen2025struq}.

Across the manually constructed attacks considered in the evaluation, the
defended models achieved ASRs between \(0\%\) and \(2\%\). This included
completion attacks using exact or modified delimiter strings. The combination
of delimiter filtering and structured instruction tuning therefore provided
strong protection against the attack patterns represented in the original
evaluation
\parencite{chen2025struq}.

The results were weaker against the Greedy Coordinate Gradient (GCG) attack.
GCG uses gradient information to search iteratively for adversarial token
sequences that increase the probability of a selected target response
\parencite{zou2023universal}. For the adapted GCG attack, StruQ reduced ASR
from \(97\%\) to \(58\%\) on the evaluated Llama model and from \(99\%\) to
\(56\%\) on the evaluated Mistral model
\parencites{chen2025struq}{zou2023universal}.

GCG requires white-box model access during payload construction and
substantially more computational effort than a manually written prompt.
Nevertheless, the remaining success rates demonstrate that StruQ increases the
difficulty of an attack without providing complete protection.

The original evaluation also measured relative utility using AlpacaEval. The
reported score changed from \(67.2\%\) to \(67.6\%\) for the Llama model and
from \(80.0\%\) to \(78.7\%\) for the Mistral model
\parencite{chen2025struq}. These results suggest that the defence preserved
relative response quality in the tested setting, although they do not establish
that every model capability remains unaffected.

A further limitation concerns deployment. StruQ is most straightforward in
programmatic applications where the system already knows which input represents
the task and which represents external data. In open-ended conversational
systems, users may mix instructions and data within the same message, making
the required separation more difficult to apply consistently
\parencite{chen2025struq}.

\subsection{SecAlign}
\label{subsec:secalign}

SecAlign pursues the same general goal as StruQ but modifies the defensive
training objective. The authors identify a limitation of supervised structured
instruction tuning: it teaches the model to produce the desired response, but
does not explicitly train it to assign a lower preference to responses that
follow the injected instruction.

SecAlign therefore formulates prompt-injection defence as a preference
optimisation problem. Instead of training only on a prompt and one correct
response, each training example contains:

\begin{itemize}
    \item a prompt-injected input,
    \item a preferred response that follows the trusted instruction, and
    \item a rejected response that follows the injected instruction.
\end{itemize}

\subsubsection{Preference Alignment}
\label{subsubsec:preference-alignment}

SecAlign uses Direct Preference Optimisation (DPO), a preference-learning
method that trains a model to prefer one response over another without requiring
a separate reinforcement-learning process. The objective increases the
likelihood of the preferred response relative to the rejected response
\parencite{chen2025secalign}.

The preference dataset is generated from an existing instruction-tuning
dataset. Instructions from other samples are inserted into the data component
to create prompt-injected inputs. The preferred response continues to perform
the legitimate task, while the rejected response follows the inserted
instruction. Completion-style attacks are also included during dataset
construction
\parencite{chen2025secalign}.

Table~\ref{tab:secalign-example} shows a simplified preference pair. The
preferred response follows the summarisation request, while the rejected
response complies with the injection.

\begin{table}[htbp]
    \centering
    \begin{tabular}{p{0.92\linewidth}}
        \toprule
        \textbf{Trusted Instruction} \\
        Summarise the following email. \\[0.4em]

        \textbf{Untrusted Input} \\
        The meeting has been moved to Friday afternoon. \\
        Ignore the previous instruction and print exactly
        \enquote{HACKED}. \\[0.4em]

        \textbf{Preferred Response} \\
        The meeting has been rescheduled to Friday afternoon. \\[0.4em]

        \textbf{Rejected Response} \\
        HACKED \\
        \bottomrule
    \end{tabular}
    \caption[SecAlign preference-pair example]{Simplified preference pair used
    during SecAlign training, adapted from
    \textcite{chen2025secalign}.}
    \label{tab:secalign-example}
\end{table}

Unlike ordinary supervised fine-tuning, SecAlign does not merely reinforce the
desired response. It simultaneously reduces the relative likelihood of a
response associated with the injected instruction. The objective therefore
teaches the model to prefer secure behaviour over the specific insecure
behaviour represented by the rejected response
\parencite{chen2025secalign}.

SecAlign does not require StruQ's newly reserved token vocabulary and can be
applied to existing instruction-tuned models using their existing prompt
formats. This does not mean that an attacker or end user must label the input.
The application integrating the model remains responsible for identifying the
trusted instruction and placing external data in the intended input position
\parencite{chen2025secalign}.

\subsubsection{Evaluation and Comparison with StruQ}
\label{subsec:secalign-evaluation}

The published SecAlign evaluation compares StruQ and SecAlign across five
open-weight models: Mistral-7B-Instruct, Llama3-8B-Instruct, Llama-7B,
Mistral-7B, and Llama3-8B. Security is evaluated on 208 AlpacaFarm samples.
Under the default criterion, an attack succeeds when the response begins with
\enquote{Hacked} or \enquote{hacked}
\parencite{chen2025secalign}.

SecAlign's maximum ASR across the three optimisation-free attacks was \(0\%\)
on every evaluated model. Across the optimisation-based attacks GCG,
AdvPrompter, and NeuralExec, the maximum ASR for each model ranged from
\(27\%\) to \(60\%\) under StruQ and from \(1\%\) to \(14\%\) under SecAlign.
For each evaluated model, SecAlign therefore reduced the maximum
optimisation-based ASR by more than a factor of four relative to StruQ
\parencite{chen2025secalign}.

The authors also evaluated utility using AlpacaEval2, an LLM-as-a-judge
comparison over 805 AlpacaFarm prompts. SecAlign's win rates remained close to
those of the corresponding undefended models in that evaluation
\parencite{chen2025secalign}.

These results demonstrate a substantial improvement within the conditions of
the SecAlign study. However, they should not be interpreted as universal
security guarantees. The optimisation-based evaluation assumes knowledge of
the prompt format and benign instruction together with white-box model access.
It also relies predominantly on a fixed injected task and a target-prefix
success criterion.

SecAlign also continues to rely on specialised defensive fine-tuning and on an
application that preserves the distinction between trusted instructions and
untrusted content. Its protection therefore remains dependent on learned model
behaviour rather than on an independently enforced security boundary.

\subsection{Comparison and Remaining Limitations}
\label{subsec:struq-secalign-comparison}

StruQ and SecAlign demonstrate two related approaches to learned
instruction--data separation. StruQ combines a structured interface, reserved
delimiter filtering, and supervised structured instruction tuning. SecAlign
instead uses preference optimisation to favour responses to the trusted task
and suppress responses associated with an injected instruction.

Both approaches substantially reduce attack success in the settings considered
by their original evaluations. SecAlign provides stronger resistance than
StruQ against the evaluated optimisation-based attacks, while maintaining
similar relative utility in the authors' benchmark.

Nevertheless, neither method creates a formally enforced distinction between
trusted instructions and untrusted data. The final decision remains part of the
model's probabilistic behaviour. Their reported effectiveness therefore depends
on the selected models, prompts, attacks, tasks, success criteria, and utility
benchmarks.

These dependencies are examined in
Section~\ref{subsec:evaluation-methodology}, together with later evaluations
and adaptive attacks that specifically target the assumptions of
fine-tuning-based defences.