% Required packages in the main preamble:
%
% \usepackage{array}
% \usepackage{booktabs}
% \usepackage{ragged2e}
% \usepackage{tabularx}

\section{Limits of the Case-Study Defences}
\label{sec:limits}

StruQ and SecAlign are meaningful improvements over undefended language models,
but their reported robustness is conditional on the design of the respective
evaluation. A low attack success rate describes a particular combination of
models, tasks, attack-construction methods, attacker access, optimisation
budgets, and success criteria. It is not a stable property that can be
transferred unchanged to another study or deployment.

This section compares the original evaluations of StruQ and SecAlign with later
work that uses broader benchmarks or attacks adapted specifically to those
defences. It also distinguishes security results from utility measurements and
documents the limitations of this review.

\subsection{Evaluation Methodology Problems}
\label{subsec:evaluation-methodology}

Attack Success Rate (ASR) is meaningful only when the event counted as a
successful attack is defined explicitly. The original StruQ and SecAlign
evaluations primarily test whether the model complies with a fixed
attacker-selected instruction.

In the StruQ evaluation, responses containing \enquote{Hacked!} or
\enquote{hacked!} are manually inspected to exclude false positives. SecAlign
uses a stricter default criterion under which the response must begin with
\enquote{Hacked} or \enquote{hacked}
\parencites{chen2025struq}{chen2025secalign}.

These criteria provide repeatable proxies for whether the model follows the
injected instruction. However, they do not directly measure confidential-data
disclosure, unauthorised tool execution, or success against open-ended attacker
objectives
\parencite{jia2025critical}.

Cross-study comparisons are further limited by prompt diversity and attacker
adaptivity. A defence evaluated against a fixed set of handcrafted injections
may not resist an attacker who knows which defence is deployed and modifies the
payload, target, or optimisation objective accordingly.

The original evaluations nevertheless provide an important baseline.
Table~\ref{tab:original-evaluations} summarises their principal security and
utility results
\parencites{chen2025struq}{chen2025secalign}.

\begin{table}[htbp]
    \centering
    \footnotesize
    \setlength{\tabcolsep}{4pt}
    \renewcommand{\arraystretch}{1.2}

    \caption{Original evaluations of StruQ and SecAlign.}
    \label{tab:original-evaluations}

    \begin{tabularx}{\linewidth}{
        @{}
        >{\RaggedRight\arraybackslash}p{2.2cm}
        >{\RaggedRight\arraybackslash}X
        >{\RaggedRight\arraybackslash}X
        >{\RaggedRight\arraybackslash}X
        @{}
    }
        \toprule

        \textbf{Defence} &
        \textbf{Security result} &
        \textbf{Utility result} &
        \textbf{Main limitation} \\

        \midrule

        StruQ
        &
        Manual attacks achieved \(0\%\)--\(2\%\) ASR. Adapted GCG still
        achieved \(58\%\) on Llama and \(56\%\) on Mistral.
        &
        AlpacaEval changed from \(67.2\%\) to \(67.6\%\) on Llama and from
        \(80.0\%\) to \(78.7\%\) on Mistral.
        &
        Only two models were evaluated, and most attacks used a fixed
        attacker-selected output.
        \\

        \addlinespace

        SecAlign
        &
        Optimisation-free attacks achieved a maximum ASR of \(0\%\).
        Optimisation-based maxima ranged from \(1\%\) to \(14\%\), compared
        with \(27\%\)--\(60\%\) for StruQ.
        &
        AlpacaEval2 win rates remained close to those of the corresponding
        undefended models in the authors' evaluation.
        &
        Results depend on the tested models, target-output criterion, and
        white-box attacker assumptions.
        \\

        \bottomrule
    \end{tabularx}
\end{table}

The original results cannot be interpreted independently of their evaluation
settings. StruQ demonstrates strong resistance to the manually constructed
attacks represented in its test suite, but its remaining GCG success rates show
that the learned separation is not absolute. SecAlign substantially improves
resistance to the optimisation-based attacks included in its evaluation, but
successful attacks remain possible.

Utility comparisons require similar care. The SecAlign authors use
AlpacaEval2, where GPT-4-Turbo judges the evaluated outputs against a
\texttt{davinci003} reference. Jia et al. instead use the corresponding
undefended Llama-3-8B-Instruct model as the relative reference and additionally
measure ground-truth task performance on OpenPromptInjection and MMLU-PI
\parencites{chen2025secalign}{jia2025critical}.

MMLU-PI denotes the prompt-injection benchmark constructed by Jia et al. from
the Massive Multitask Language Understanding benchmark. Their study therefore
uses different reference responses, model variants, tasks, and success
definitions from the original SecAlign evaluation.

In the Jia evaluation, StruQ and SecAlign obtain relative win rates of
\(21.60\%\) and \(36.42\%\), respectively, against the undefended
instruction-tuned model. Their absolute OpenPromptInjection/MMLU-PI scores are
\(0.54/0.35\) for StruQ and \(0.48/0.34\) for SecAlign, compared with
\(0.65/0.45\) for the undefended model
\parencite{jia2025critical}.

These results demonstrate utility loss under the model, reference, and tasks
used by Jia et al. They do not retroactively invalidate the different relative
comparison reported in the original SecAlign study. Instead, the disagreement
illustrates that utility depends on the chosen benchmark, judging method,
reference model, and task distribution.

\subsection{Adaptive Attacks}
\label{subsec:adaptive-attacks}

An adaptive attack targets assumptions, components, or training procedures
specific to the defence being evaluated. Such attacks are particularly
informative when assessing worst-case robustness, although their access
requirements may be less representative of attacks against closed commercial
deployments.

Later evaluations use broader benchmarks or attacks designed specifically
around the defended models. Their principal results are summarised in
Table~\ref{tab:adaptive-evaluations}
\parencites{jia2025critical}
             {yang2025checkpointgcg}
             {pandya2025astra}.

\begin{table}[htbp]
    \centering
    \footnotesize
    \setlength{\tabcolsep}{4pt}
    \renewcommand{\arraystretch}{1.2}

    \caption{Later and defence-adaptive evaluations.}
    \label{tab:adaptive-evaluations}

    \begin{tabularx}{\linewidth}{
        @{}
        >{\RaggedRight\arraybackslash}p{2.5cm}
        >{\RaggedRight\arraybackslash}p{3.0cm}
        >{\RaggedRight\arraybackslash}X
        >{\RaggedRight\arraybackslash}X
        @{}
    }
        \toprule

        \textbf{Evaluation} &
        \textbf{Attacker access} &
        \textbf{Principal result} &
        \textbf{Interpretation} \\

        \midrule

        Jia et al.
        &
        Broader attacks and adaptive white-box GCG with knowledge of the
        delimiter defence.
        &
        Adaptive GCG achieved an ASV of \(1.00\) against StruQ on both tested
        benchmarks. SecAlign also showed reduced security and utility.
        &
        Robustness changes substantially when the benchmark, success rule,
        model, and utility reference are changed.
        \\

        \addlinespace

        Checkpoint-GCG
        &
        White-box access to intermediate fine-tuning checkpoints; limited
        black-box transfer evaluation.
        &
        Exact-input ASRs reached \(84\%\)--\(96\%\). Universal attacks reached
        \(75.3\%\)--\(89.9\%\).
        &
        Fine-tuning obstructed ordinary GCG but did not remove the underlying
        attack surface. The audit requires unusually strong access.
        \\

        \addlinespace

        ASTRA and ASTRA++
        &
        White-box access to weights, gradients, attention information, and
        defence details.
        &
        ASTRA achieved \(72.5\%\) and \(82.5\%\) ASR. ASTRA++ reached
        \(96\%\) against SecAlign and \(87\%\) against SecAlign++.
        &
        Attention-aware attacks bypassed the evaluated defences, although the
        access assumptions do not represent every closed deployment.
        \\

        \bottomrule
    \end{tabularx}
\end{table}

The evaluations in Table~\ref{tab:adaptive-evaluations} do not form a common
ranking. They use different models, datasets, attack-construction methods,
optimisation budgets, attacker capabilities, and success criteria. Results
should therefore be compared primarily within each study rather than directly
across the reported percentages.

Jia et al. use a task-specific Attack Success Value (ASV) rather than only a
binary target-output ASR. In their adaptive GCG evaluation, StruQ reaches an
ASV of \(1.00\) on both OpenPromptInjection and MMLU-PI. SecAlign reaches
\(0.46\) on OpenPromptInjection and \(0.88\) on MMLU-PI
\parencite{jia2025critical}.

The results indicate that StruQ and SecAlign are less robust when tested using
broader tasks and attacks adapted to their defensive mechanisms. However, the
models, datasets, utility references, and success measures differ from those
used in the original evaluations.

\subsubsection{Checkpoint-GCG}

Checkpoint-GCG attacks a sequence of intermediate fine-tuning checkpoints.
GCG is first applied to an early checkpoint. The resulting adversarial suffix
is then used to initialise optimisation against the next selected checkpoint.
This process continues through the sequence, using each suffix as a stepping
stone towards the final defended model
\parencite{yang2025checkpointgcg}.

In the exact-input auditing setting, the attack assumes access to the complete
model input and to intermediate fine-tuning checkpoints. Evaluated on 50
randomly selected AlpacaFarm samples, Checkpoint-GCG achieved ASRs of \(88\%\),
\(96\%\), and \(84\%\) against SecAlign-defended Llama, Mistral, and Qwen
models
\parencite{yang2025checkpointgcg}.

Universal suffixes constructed from ten AlpacaFarm prompts achieved
\(75.3\%\), \(89.9\%\), and \(78.3\%\) ASR on 198 held-out prompts,
respectively. A suffix constructed using the Llama model additionally achieved
\(63.9\%\) through black-box queries against the related
Meta-SecAlign-8B target
\parencite{yang2025checkpointgcg}.

The latter result provides evidence of transfer for one closely related model
and defence pair. The reported Mistral- and Qwen-derived suffixes did not
transfer successfully in the corresponding experiments.

Checkpoint-GCG therefore demonstrates that resistance to ordinary GCG does not
establish resistance to an attacker who can exploit the fine-tuning process.
However, checkpoint access represents a strong audit assumption and is not
available to every real-world attacker.

\subsubsection{ASTRA and ASTRA++}

ASTRA uses a two-stage white-box optimisation process. It first optimises an
attention-based objective to produce an initial adversarial sequence. This
sequence is then refined using a GCG-style objective targeting the selected
model output
\parencite{pandya2025astra}.

The attention-based stage differs from ordinary GCG because it considers how
the model allocates attention between the trusted instruction and the injected
content. It attempts to construct an adversarial sequence that changes this
relationship before optimising directly for the target output.

In the strong-context-knowledge evaluation, the authors completed 40
AlpacaFarm examples using 500 optimisation iterations. At a 35-token budget,
consisting of 15 prefix tokens and 20 suffix tokens, ASTRA achieved
\(72.5\%\) ASR against SecAlign-defended Llama-3-8B-Instruct and
\(82.5\%\) against SecAlign-defended Mistral-7B-Instruct-v0.1
\parencite{pandya2025astra}.

ASTRA++ evaluates a weaker context-knowledge setting in which the preceding
context is not available during optimisation, while white-box model and defence
access remain available. Using ten training prompts and 198 held-out prompts,
the 35-token configuration achieved a mean ASR of \(96\%\), with a standard
deviation of five percentage points, against SecAlign-defended Llama.

Against SecAlign++, where the trusted instruction occupies the system role,
ASTRA++ achieved a mean ASR of \(87\%\), with a standard deviation of twelve
percentage points
\parencite{pandya2025astra}.

These results demonstrate that attention-guided attacks can bypass the tested
defences under the stated white-box conditions. They do not establish the same
success against closed APIs, dynamically changing system contexts, arbitrary
attacker objectives, or deployments with additional application-level
controls.

Checkpoint-GCG and ASTRA consequently support a narrower conclusion than the
claim that every model-level defence must fail. They demonstrate that StruQ and
SecAlign do not create a deterministic or independently enforced trust boundary
under the evaluated adaptive conditions.

\subsection{Security and Utility Trade-Offs}
\label{subsec:security-utility}

Defensive training must reduce malicious instruction following without
preventing legitimate analysis of instructional language. A contract-review
task, for example, may require the model to quote, classify, or reason about
commands without executing them.

A model trained to distrust every imperative sentence in the data component
could therefore reject or mishandle legitimate inputs. Conversely, a model that
preserves unrestricted instruction-following behaviour may retain a larger
prompt-injection attack surface.

The original StruQ and SecAlign studies report only small changes on their
relative LLM-judged utility benchmarks. The later Jia evaluation reports lower
relative and absolute utility for the released defended Llama variants under a
different reference and different ground-truth tasks.

The defensible synthesis is not that one evaluation invalidates the other.
Instead, utility depends on the evaluated model, task distribution, reference
response, judging method, and benchmark
\parencites{chen2025struq}{chen2025secalign}{jia2025critical}.

Security and utility must therefore be reported together. A defence that
refuses most inputs may record a low attack success rate while being unusable
in practice. A defence that preserves broad instruction-following capabilities
may remain more useful while allowing a larger set of adversarial instructions
to influence its behaviour.

Detector-based defences require additional measurements. Their evaluations
should report false-positive and false-negative rates rather than relying only
on aggregate accuracy or area under the curve
\parencite{jia2025critical}.

\subsection{Limitations of This Review}
\label{subsec:review-limitations}

This paper is a selective narrative review rather than a systematic literature
survey. Its empirical analysis concentrates on StruQ and SecAlign and on later
work that re-evaluates or adapts attacks to those defences. No original
experiments or independent replications were conducted.

Detector-based and vendor-specific defences receive only limited treatment.
Closed providers may not disclose the model versions, prompts, filters,
training methods, or configuration changes required for reproducible
evaluation. Much of the evidence concerning the newest adaptive attacks also
remains available only as recent preprint research.

Comparison is restricted by differences in models, prompts, tasks, attacker
budgets, attacker knowledge, defence knowledge, context availability, success
definitions, and utility baselines. Changing preprint versions and updates to
deployed models further limit how long individual results remain applicable.

Target-output ASR is also an incomplete proxy for practical harm. Printing a
selected word demonstrates that the model followed an injected instruction,
but does not directly measure confidential-data disclosure, unauthorised tool
use, financial impact, or persistence across interactions.

The review is limited to text-based prompt injection. It does not evaluate
multimodal injection, persistent memory attacks, denial of service, malicious
insiders, provider compromise, or training-data poisoning.

Architectural controls are themselves conditional. Their guarantees depend on
correct trust labels, sufficiently complete policies, trusted interpreters or
security monitors, correctly configured tool wrappers, and an uncompromised
trusted computing base
\parencites{wu2024systemlevel}{debenedetti2025defeating}.

The conclusions therefore apply to the sources, models, and assumptions
examined in this paper rather than to every possible prompt-injection defence.

\subsection{Implications of the Evidence}
\label{subsec:open-problem}

The reviewed evidence supports a tiered answer to the research question.

First, learned instruction--data separation substantially reduces the success
of the handcrafted attacks considered in the original StruQ and SecAlign
evaluations.

Second, results against standard optimisation-based attacks are mixed but still
show meaningful improvement. SecAlign, in particular, performs substantially
better than StruQ against the optimisation-based attacks included in its
published evaluation.

Third, later defence-adaptive white-box attacks demonstrate that this
robustness is not an independently enforced security guarantee. Access to
fine-tuning checkpoints, gradients, attention information, or defence details
allows attackers to target assumptions that are not represented by the
original attack suites.

Finally, the extent to which these audit results transfer to closed,
dynamically configured production systems remains unresolved. White-box
bypasses establish that the model-level separation is not absolute, but they do
not determine the success rate of every attacker against every deployment.

Separating instructions and data should therefore be treated as a
risk-reduction mechanism rather than as complete prevention. Model-level
protection remains useful, but secure deployments must assume that it may fail
and constrain the resulting information flows and actions outside the model.