\section{Benchmarks and Metrics}
\label{sec:benchmarks}

Having examined various defence mechanisms, the focus must shift to measuring
their performance in terms of both effectiveness and usability. A principled,
modern approach to evaluation is critical, as many early claims of robustness
relied largely on overly narrow testing conditions~\parencite{jia2025critical}.

\subsection{Metrics for Utility and Reliability}

The primary challenge in defending LLMs is maintaining their general-purpose
capabilities and usability while enforcing security. Thus, researchers often use
relative utility, measured via win rates on benchmarks such as AlpacaEval
~\parencite{jia2025critical}. In this setup, a strong model like GPT-4 judges
whether a defended model's output is better or worse than a reference model
~\parencite{chen2025struq}.

However, relative win rates can be misleading; recent studies show that while
StruQ and SecAlign maintained high win rates, they suffered significant drops
in absolute utility when measured on ground-truth dataset-based benchmarks such
as MMLU and OpenPromptInjection~\parencite{jia2025critical}.

For detection-based defences, utility is further quantified by the False
Positive Rate (FPR). A high FPR, meaning harmless prompts are misclassified as
attacks, renders a defence impractical by causing user fatigue and the rejection
of legitimate queries~\parencite{jia2025critical}.

\subsection{Security Effectiveness Metrics}

The primary standard for security evaluation is the Attack Success Rate (ASR),
defined as the percentage of cases where the LLM follows an injected instruction
instead of the developer's prompt~\parencite{zou2023universal}.

For complex tasks where output is not binary, such as summarisation,
researchers utilise the Attack Success Value (ASV). ASV employs similarity
scores to quantify the degree to which a model's output aligns with the
adversary's intent~\parencite{jia2025critical}.

Additionally, defences based on preference optimisation, such as SecAlign,
utilise the log-probability margin between desirable and undesirable outputs to
measure the strength of the model's internal rejection of malicious instructions
~\parencite{chen2025secalign}.

\subsection{Standardised Benchmark Suites}

Recent research has moved toward large-scale, standardised suites to prevent
over-optimisation for specific attacks. OpenPromptInjection provides a diverse
set of prompt-response pairs across seven fundamental Natural Language
Processing (NLP) tasks. InjecAgent and AgentDojo are two benchmark suites which evaluate models in dynamic, tool-integrated environments where successful prompt injections do not directly cause harm or data exfiltration within the benchmark environment, even if the injection succeeds \parencites{wu2024systemlevel}{zhong2025rtbas}.

Crucially, evaluation must include adaptive attacks such as ASTRA, a specialised
attack against SecAlign and StruQ~\parencite{pandya2025astra}, which targets
the attention mechanisms of the model. Defences often appear more robust on
static datasets than they are in realistic adversarial settings
~\parencite{jia2025critical}.

Overall, the evaluation of prompt injection defences requires balancing
security and utility. Metrics such as ASR, ASV, win rates, and FPR capture
different aspects of this trade-off, while standardised benchmarks provide a
more realistic basis for comparison across approaches. Applying these methods
has revealed important weaknesses in several established defences, including
StruQ and SecAlign.

These metrics and benchmark suites therefore serve a dual purpose. They not
only enable comparisons between different defence mechanisms, but also expose
the gap between reported robustness and practical security under realistic
adversarial conditions.

Understanding these limitations is essential when assessing the true
effectiveness of current approaches. The weaknesses uncovered through such
evaluations motivate a closer examination of where existing defences break
down, which is the focus of the following section.

