\section{Practical Implications and Deployment Considerations}
\label{sec:praxis}

\subsection{Prompt Injection as a System-Level Risk}

The impact of prompt injection depends on the data and authority exposed by the
surrounding application. A read-only summarisation system may return a
manipulated answer, whereas an email assistant or agent may access private
resources, send messages, or invoke external tools. The same model-level failure
can therefore have different consequences under different permission models
\parencite{debenedetti2024agentdojo}.

The researcher-reported Slack AI demonstration illustrates this dependence. A
malicious instruction in a public channel influenced a response that also used
private-channel data. The generated link encoded the protected content, but the
victim still had to click it for exfiltration to occur
\parencite{promptarmor2024slack}. The chain depended on retrieval scope, the
model's output, the link destination, and human interaction; no one component
alone describes the complete risk.

\subsection{Deployment Principles}

\paragraph{Preserve provenance and authority.}

External documents, direct user input, tool outputs, summaries derived from
them, and records stored in memory should retain their trust status throughout
the application. This supports instruction integrity by preventing data from
silently becoming an authorised control instruction. A label or delimiter that
appears only in the prompt remains model-mediated. It becomes an externally
enforced control only when the application preserves the label and a policy
component uses it when authorising information flows or actions
\parencites{hines2024spotlighting}{wu2024systemlevel}.

This control assumes that sources are labelled correctly and that transformations
such as retrieval, summarisation, and memory updates preserve those labels.
Incorrect or missing labels remain a residual risk.

\paragraph{Restrict tools, credentials, and parameters.}

Least privilege protects confidentiality and action integrity by limiting what a
successful injection can access or change. Credentials should be short-lived and
scoped to the current task. Read authority should be separated from write, send,
or delete authority, and tools should be restricted to intended destinations and
recipients
\parencites{debenedetti2024agentdojo}{debenedetti2025defeating}
{owasp2025llm}.

An authorisation gateway outside the model should validate the requested
operation, destination, recipient, schema, and security-relevant parameters. A
model-generated argument should not acquire authority merely because it is
syntactically valid. These checks assume that the policy captures the relevant
constraints and that tool wrappers cannot be bypassed. They may still permit an
undesired action that falls within an overly broad authorised set.

\paragraph{Contain execution.}

Sandboxing protects action integrity by restricting file, network, code-execution,
and secret access even when model output is malicious. It is enforced by the
execution environment rather than by the LLM. A sandbox does not prevent prompt
injection, and authorised network or tool bridges remain part of the attack
surface. Its purpose is containment, not proof that the generated action is
benign.

\paragraph{Use independent approval for high-impact actions.}

Sending a message, modifying a record, disclosing protected data, or making an
irreversible transaction should require an independent policy decision or
explicit human approval. The approval interface should display the exact action,
recipient, destination, disclosed data, and expected consequence independently
of the model's explanation. This protects action integrity and confidentiality
only if the reviewer can understand the decision and is not encouraged to
approve routinely. Human error, misleading presentation, and social engineering
remain residual risks.

\subsection{Testing, Monitoring, and Incident Response}

Monitoring and detection are separate from prevention. Protected logs should
record relevant input provenance, retrieved content, model output, proposed tool
calls, policy decisions, approvals, credential identifiers and scopes used, and
external destinations---never credential secrets. Because those records may
contain sensitive prompts and outputs, they require access control, data
minimisation, retention limits, and integrity protection. Such records support
detection and investigation, but they do not undo a disclosure or prevent the
initial model failure
\parencite{owasp2025llm}.

Pre-deployment and regression testing should use the application's real trust
boundaries and permissions. It should include direct and indirect payloads,
multi-step tool interactions, diverse target and injected tasks, and attacks
adapted to the deployed defence. White-box audits are useful for testing the
limits of open-weight defences, but should be reported separately from realistic
query-only attack paths
\parencites{debenedetti2024agentdojo}{jia2025critical}
{yang2025checkpointgcg}{pandya2025astra}.

An incident plan is required when an injection succeeds. The operator should be
able to contain the affected workflow, revoke active or exposed credentials,
quarantine malicious content, preserve and inspect logs, determine which data
and actions were affected, and add the incident to regression tests. Malicious
content may also need to be removed from retrieval stores or memory before the
workflow is resumed. These recovery steps reduce continued exposure; they cannot
guarantee recovery of already disclosed information.

\subsection{Trade-Offs and Outlook}

StruQ and SecAlign require model adaptation and, for structured interfaces,
corresponding application support. Their integration cost is application-specific,
and their protection remains model-mediated and probabilistic. Information-flow
control and CaMeL-style architectures can provide stronger, externally
enforceable guarantees under explicit policy and enforcement assumptions
\parencites{wu2024systemlevel}{debenedetti2025defeating}.

Those guarantees depend on correct data labelling, sufficiently complete
policies, trusted interpreters or security monitors, correct tool wrappers, and
an uncompromised trusted computing base. They do not ensure that the language
model will never follow an injected instruction. Instead, they can constrain
whether model behaviour becomes an authorised information flow or external
action.

These controls also impose operational costs: policy gates and sandboxes require
integration and maintenance, approvals add latency and reduce automation, and
monitoring creates storage, access-control, and privacy obligations.

Deployment decisions must balance engineering cost, task utility, policy
coverage, and the consequences of both model and enforcement failures. Where
external controls cannot reduce the remaining impact to an acceptable level,
the operator should document and accept that residual risk explicitly or avoid
automating the action. Defence in depth is therefore not a claim that many weak
layers become infallible; it is a strategy for combining behavioural mitigation,
independent prevention, containment, detection, and recovery.
