\section{Limits of Current Defences}
\label{sec:limits}

\subsection{Evaluation Methodology Problems}

A recurring problem in prompt injection defence research is that reported
robustness numbers often do not reflect realistic adversarial conditions.
\citeauthor{jia2025critical} argue that the field has systematically evaluated
defences under settings that are too narrow to support the conclusions drawn
from them~\cite{jia2025critical}.

The most common issue is attack diversity. Many evaluations rely primarily on
heuristic attacks --- manually written strings such as ``ignore previous
instructions'' --- as their primary or sole measure of robustness. These serve
as a useful baseline, but they do not represent an attacker who actively adapts
their strategy to the specific defence being tested. A defence that blocks every
heuristic attack in the test set may still be trivially bypassed by an attacker
who knows how it works.

A second issue is prompt diversity. Several defences, including StruQ and
SecAlign in their original evaluations, measure attack success rate against a
single injected prompt: ``Print exactly `Hacked!'\,''. As \citeauthor{jia2025critical}
demonstrate, this significantly overstates robustness. When evaluated across
diverse target and injected prompt combinations drawn from OpenPromptInjection
and MMLU-PI, GCG achieves ASV exceeding 0.80 against StruQ-defended models and
0.72 against SecAlign-defended models --- results that stand in sharp contrast
to the near-zero rates reported in the original papers~\cite{jia2025critical}.

A third issue concerns utility measurement. Several defences report utility
preservation using win rate on AlpacaFarm, comparing the defended model against
a reference LLM. \citeauthor{jia2025critical} show that this metric is
insufficient: when using Llama-3-8B-Instruct as a more appropriate reference,
StruQ achieves a win rate of only 21.6\% and SecAlign 36.4\%. On absolute
utility benchmarks, both models show measurable capability degradation compared
to the undefended baseline~\cite{jia2025critical}. A model can appear secure
and utility-preserving under one evaluation methodology while performing
substantially worse under a more rigorous one.

\subsection{Adaptive Attacks}

The clearest evidence that current defences are less robust than reported comes
from adaptive attacks --- attacks specifically designed to target the defence
mechanism rather than the underlying model.

\citeauthor{yang2025checkpointgcg} introduce Checkpoint-GCG, which exploits
the incremental nature of fine-tuning to find adversarial suffixes that
transfer to the final defended model. The key observation is that intermediate
training checkpoints are substantially more vulnerable than the final model.
By running GCG against each checkpoint in sequence --- using the successful
suffix from one checkpoint to initialise the next --- the attack maintains
traction across the training process. Against SecAlign-defended
Llama-3-8B-Instruct, standard GCG achieves only 6\% ASR. Checkpoint-GCG
achieves 88\%~\cite{yang2025checkpointgcg}. For universal suffixes that
generalise across unseen inputs, Checkpoint-GCG reaches 75.3\% success on
held-out samples and transfers to Meta-SecAlign-8B at 63.9\% in a black-box
setting --- without access to the target model's
weights~\cite{yang2025checkpointgcg}.

This result has an important implication. The apparent robustness of SecAlign
against standard GCG is not evidence that the underlying vulnerability has been
eliminated. It reflects the fact that GCG's gradient signal becomes a poor
optimisation guide once the loss landscape is reshaped by fine-tuning.
Checkpoint-GCG sidesteps this by approaching the defended model gradually,
from a direction where the landscape is still navigable.

\citeauthor{pandya2025astra} demonstrate a related but architecturally distinct
failure mode. Their attack, ASTRA, targets the attention matrices of the model
directly rather than optimising for output token probabilities. The key insight
is that if the model's attention can be redirected toward the attacker's
injected content, the model will follow that content regardless of what the
fine-tuning defence has trained it to do. Against SecAlign-defended
Mistral-7B-Instruct, ASTRA substantially outperforms standard GCG at the same
token budget~\cite{pandya2025astra}. The universal variant, ASTRA++, reaches
96\% success on unseen prompts against SecAlign-defended Llama-3-8B-Instruct
--- operating under the weak knowledge assumption where the attacker has no
access to the surrounding context~\cite{pandya2025astra}.

\citeauthor{pandya2025astra} also provide an explanation for why GCG
consistently underperforms as an evaluation tool for these defences. Through
controlled experiments comparing guided and unguided search, they show that
GCG's gradient-based candidate selection does not significantly outperform
random token substitution against defended models. The gradient of the target
log-probability provides a poor optimisation signal once the defence has
reshaped the loss landscape. This means that GCG was never an appropriate
evaluation mechanism for fine-tuning-based defences --- not because the
defences are strong, but because the attack is not suited to the
task~\cite{pandya2025astra}.

\subsection{Security and Utility Trade-offs}

Robustness does not come for free. Every defence mechanism discussed in this
paper imposes some cost on model utility, and the magnitude of that cost is
frequently underreported.

Detection-based approaches introduce false positives. The ProtectAI BERT
classifier evaluated in AgentDojo reduces targeted ASR to 7.95\%, but also
reduces benign utility to 41.49\% --- the largest utility drop of any defence
tested~\cite{debenedetti2024agentdojo}. PromptGuard also illustrates the risk
of over-detection: while it performs well on aggregate detection metrics,
\citeauthor{jia2025critical} report a high false positive rate on clean
inputs~\cite{jia2025critical}. A detector with this kind of false positive
behaviour is not deployable in practice.

Fine-tuning defences impose subtler but still measurable costs. As discussed
above, StruQ and SecAlign both show absolute utility degradation when evaluated
against appropriate baselines~\cite{jia2025critical}. Tool isolation in
AgentDojo reduces ASR to 6.84\% but fails entirely in 17\% of test cases where
the tools required to complete the legitimate task overlap with those required
to carry out the attack~\cite{debenedetti2024agentdojo}.

The deeper issue is that robustness and instruction-following capability are in
tension by design. A model trained to be more sceptical of instructions
embedded in data will sometimes be too sceptical --- treating legitimate
imperative sentences in a document as potential injections. SecAlign partially
mitigates this: its evaluation shows that the defended model is actually better
at processing imperative sentences in data as data, rather than as
instructions, compared to the undefended baseline~\cite{chen2025secalign}. But
this does not fully resolve the tension, and it does not hold uniformly across
all deployment contexts.

\subsection{Architectural Limitations}

Fine-tuning defences operate within a fundamental constraint: all content still
enters the same context window and is processed by the same model. The
separation between trusted instructions and untrusted data that StruQ and
SecAlign attempt to enforce is semantic, not structural. It is a learned
behaviour, not a hard boundary. A sufficiently capable attacker who understands
how the model was trained can search for inputs that make the model behave as
if that boundary does not exist --- which is precisely what Checkpoint-GCG and
ASTRA demonstrate~\cite{yang2025checkpointgcg, pandya2025astra}.

System-level approaches attempt to address this at the architectural level.
\citeauthor{wu2024systemlevel} apply information flow control principles to
enforce a hard separation between trusted planning components and untrusted
external data, ensuring that injected content cannot influence the planning
process regardless of its linguistic content~\cite{wu2024systemlevel}.
\citeauthor{zhong2025rtbas} take a related approach in RTBAS, using selective
taint tracking and dependency screening to contain the influence of untrusted
tool outputs~\cite{zhong2025rtbas}. These approaches provide stronger formal
guarantees than model-level defences, but they require significant changes to
system architecture and reduce the flexibility that makes LLM agents useful in
the first place.

The tension between formal guarantees and practical deployability has not been
resolved. A defence that is architecturally secure but requires a complete
redesign of the application pipeline will not be adopted. A defence that is
easy to deploy but relies on a model's learned behaviour to hold under pressure
provides weaker guarantees than its evaluation numbers suggest.

\subsection{Why Prompt Injection Remains an Open Problem}

The evidence reviewed in this section points to a consistent pattern. Defences
that appear robust under standard evaluation conditions tend to weaken
significantly when the evaluation methodology is made more rigorous ---
whether by diversifying the attack set, introducing adaptive attacks, or
measuring utility more carefully. This mirrors a well-established dynamic in
adversarial machine learning, where defences evaluated against a fixed attacker
are routinely broken once a stronger adversary is
applied~\cite{jia2025critical}.

The deeper issue is structural. Prompt injection exploits the core capability
of LLMs --- their ability to follow natural language instructions --- and turns
it against the system. As long as models process instructions and data in the
same context window without structural separation, the attack surface remains.
Fine-tuning can raise the cost of a successful attack, but it cannot eliminate
the underlying condition. Detection can intercept known attack patterns, but
adaptive attackers learn to evade detectors as quickly as detectors are
trained~\cite{jia2025critical, yang2025checkpointgcg}.

As LLMs gain access to external tools, APIs, memory systems, and autonomous
decision-making capabilities, the stakes of a successful injection continue to
rise. The problem is not a temporary implementation flaw awaiting a patch.
No universally reliable solution currently exists, and it remains an active
area of security research.
