\section{Practical Implications and Deployment Considerations}
\label{sec:praxis}

\subsection{Prompt Injection as a System-Level Risk}

A key conclusion from the research surveyed in this paper is that prompt
injection should not be treated as an isolated model weakness. In most
real-world deployments, the LLM is one component within a larger pipeline that
includes retrieval systems, external APIs, tool execution frameworks, memory
systems, and user-facing interfaces. A successful injection does not stop at
the model --- it propagates into whatever the model has access to.

This reframes the threat significantly. The relevant question is not simply
whether the model can be manipulated, but what capabilities the surrounding
system exposes once manipulation occurs. A model with read-only access to a
document store poses a different risk profile than one with the ability to send
emails, execute code, or authorise financial transactions. As AgentDojo
demonstrates, prompt injection becomes substantially more dangerous the more
autonomously capable the system is~\cite{debenedetti2024agentdojo}. In the
Slack AI incident documented by \citeauthor{chen2025secalign}, a single
injected instruction in a public channel was sufficient to exfiltrate data from
private conversations --- not because the model was poorly designed, but
because it had access to data the attacker wanted~\cite{chen2025secalign}.

The P2SQL attack family illustrates a further dimension of this problem. When
an LLM generates database queries from natural language, a malicious prompt can
manipulate the generated SQL, turning a prompt injection into a classical SQL
injection attack against the backend~\cite{pedro2025p2sql}. The LLM becomes a
bridge between natural language and traditional injection vulnerabilities ---
a connection that most threat models do not account for.

\subsection{Defence in Depth}

Given that no single defence mechanism eliminates the attack surface, the
practical security literature increasingly converges on a layered approach.
The goal is not to prevent every successful injection --- which current evidence
suggests is not achievable --- but to ensure that a successful injection cannot
easily escalate into a critical incident.

A layered deployment might combine detection-based filtering at the input
boundary, structured separation of instructions and data at the model level,
restricted tool permissions that limit what the model can do even if
manipulated, sandboxed execution environments that contain the consequences of
any action, and human approval requirements for high-risk or irreversible
operations. No single layer is sufficient, but each reduces the probability
that an attacker can traverse the full chain from injection to impact.

This mirrors established practice in traditional software security, where
defence in depth has long been the standard response to vulnerabilities that
cannot be fully eliminated at any single layer. The same logic applies here.
SecAlign's authors explicitly acknowledge this framing, recommending that their
defence be combined with detection systems and system-level controls rather
than deployed as a standalone solution~\cite{chen2025secalign}.

\subsection{Human Oversight and Operational Controls}

Current research consistently shows that LLM behaviour under adversarial
conditions remains difficult to predict. Fully autonomous systems that execute
sensitive actions without human review therefore introduce risk that
purely technical defences cannot fully mitigate.

Practical operational controls do not solve prompt injection directly, but they
limit its consequences. Requiring explicit user confirmation before sensitive
or irreversible actions --- sending an email, making a purchase, modifying a
file --- means that a successful injection must also deceive the user before
it can cause harm. Restricting which tools an agent can access by default
reduces the attack surface available to an injected instruction. Logging model
outputs and tool calls creates an audit trail that supports incident detection
and response.

AgentDojo's tool filter defence illustrates how even a simple form of
operational control can be effective: by requiring the model to declare which
tools it needs before processing any untrusted data, the defence reduces
targeted ASR to 6.84\%~\cite{debenedetti2024agentdojo}. The limitation ---
that it fails when user and attacker require the same tools --- also illustrates
why operational controls alone are insufficient and need to be combined with
other layers.

\subsection{Deployment Trade-offs}

Secure LLM deployment requires navigating genuine trade-offs between security,
usability, and operational complexity. These trade-offs do not have clean
solutions and are likely to remain a source of tension in commercial deployments.

Architectural approaches based on information flow control provide the strongest
formal guarantees currently available, but require substantial redesign of
existing application pipelines~\cite{wu2024systemlevel, zhong2025rtbas}. An
organisation that has already built an agentic system around a standard LLM API
cannot easily retrofit structural isolation. Fine-tuning approaches such as
StruQ and SecAlign integrate more naturally into existing workflows, but their
robustness is contingent on the attacker not having access to the model weights
or training checkpoints --- an assumption that does not hold for open-weight
models deployed in adversarial environments~\cite{yang2025checkpointgcg}.

Detection systems are the most operationally lightweight option, but as
discussed in Section~\ref{sec:limits}, they impose utility costs through false
positives and remain bypassable through adaptive
attacks~\cite{jia2025critical}. The PromptGuard example, where high AUC
coexists with a very high false positive rate on clean inputs, illustrates why
aggregate metrics can be misleading when evaluating detection-based
defences~\cite{jia2025critical}.

In commercial environments, deployment speed and usability pressures frequently
conflict with conservative security design. This is not a failure of engineering
judgement --- it reflects genuine uncertainty about which threats will
materialise and what level of friction users will tolerate. What the research
reviewed in this paper does establish clearly is that the cost of underestimating
the threat is likely to increase as LLM systems gain more autonomy and access
to more sensitive operations.

\subsection{Future Directions}

Several directions emerge naturally from the limitations identified in this
paper. At the technical level, the most promising avenue is stronger
architectural isolation --- moving away from relying on model behaviour to
enforce trust boundaries and toward structural mechanisms that make those
boundaries enforceable regardless of what the model does. Work in this
direction, such as \citeauthor{wu2024systemlevel}'s information flow control
approach and \citeauthor{debenedetti2025defeating}'s design-level
guarantees~\cite{wu2024systemlevel, debenedetti2025defeating}, represents a
fundamentally different research programme than fine-tuning defences, and one
that may be necessary if the adaptive attack results in Section~\ref{sec:limits}
are taken seriously.

At the evaluation level, the field needs methodology that takes adaptive
attackers as the baseline rather than the exception. Benchmarks such as
AgentDojo and FSPIB are steps in the right direction because they evaluate
systems under more realistic conditions~\cite{debenedetti2024agentdojo,
fspib2025}, but standardised evaluation against adaptive attacks with
consistent utility metrics remains an open problem. The critique by
\citeauthor{jia2025critical} of current evaluation practice should be treated
as a call for methodological reform, not just a criticism of individual
papers~\cite{jia2025critical}.

One further complication deserves mention. More capable models are not
necessarily more secure. AgentDojo's inverse scaling result --- that more
capable models tend to be easier to attack because they are better at executing
whatever instructions they receive --- suggests that the relationship between
model capability and security is not straightforward~\cite{debenedetti2024agentdojo}.
Improvements in general model quality do not automatically translate into
improvements in prompt injection robustness. This is a consideration that
organisations deploying frontier models should factor into their threat
assessments.
